Solutions to Integer Programming from Quantum Annealing

### 2020-03-11 paper/main.tex
Updated initial QUBO to factored out T
Synced presentation with paper
Re-included transpose adjacency for mds
Added classical section

\subsection{Integer Programming Mapping to QUBO}
\label{sec:results:ilp}
The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
\chi^2 = & \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu + p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],\\
\Psi = & \begin{pmatrix}
\Psi^x\\
\Psi^s
\end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization. The objective function can be represented as a QUBO Hamiltonian
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
 Q_{00} & Q_{01}\\
Q_{10} & {T^{s}}^T Q_{11} T^s
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + pb^2,\\
\equiv & \Psi^T Q \Psi + C,
\label{eq:matrix_form}
\end{align}
where
\begin{align}
Q_{00} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x + \mathrm{Diag}_{\psi^x}(c) T^x,\\
Q_{01} = & Q_{10}^T = p {T^{x}}^T A^T T^s,\\
Q_{11} = & p\mathrm{Diag}_{\psi^s}\left(1 + b^T + b\right).
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

@@

Given a graph $G(E,V)$, defined by the set of $V$ vertices and $E$ edges, and a subset of vertices $D \subseteq V$. If the set of all vertices not in $D$ is adjacent to at least one vertex in $D$, then $D$ is a dominating set. This is equivalent to requiring the set of nearest-neighbor vertices of $D$ (exclusive) and $D$ cover all vertices $N(D) \cup D = V$ (an example is given by Fig.~\ref{fig:dominating_sets}a). The set $D$ is a minimal dominating set if there is no proper subset of $D$ that is a dominating set, {\it{i.e.}}, the removal of any vertex in $D$ results in $N(D) \cup D  \neq V$. An example is given by Fig.~\ref{fig:dominating_sets}b. The domination number of $D$ is given by the cardinality of $|D| \equiv \overline{\overline{D}}$. The minimum dominating set is defined by $D$ with the smallest domination number. Fig.~\ref{fig:dominating_sets}c shows an example of the minimum dominating set of $G(V, E)$ and is different from the minimal dominating set. We emphasize that while the maximum independent set is always a minimal dominating set as exemplified by Fig.~\ref{fig:dominating_sets}b, the minimum dominating set in general can have a smaller domination number. As a result, the solution to the dominating set problem can not be obtained by solving for the maximum independent set, a problem that is well studied for quantum annealers~\cite{}.

@@

@@

The target Hamiltonian can be written in the matrix form
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
\mathbbm{1} + p\left[J^2 + 2 J - \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] & - p(\mathbbm{1}+J)T^s\\
- p{T^s}^T(\mathbbm{1}+J)& p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + p
\label{eq:matrix_form}
\end{align}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

>>>>

\subsection{Integer Programming Mapping to QUBO}
\label{sec:results:ilp}
The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
\chi^2 = & \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu + p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],\\
\Psi = & \begin{pmatrix}
\Psi^x\\
\Psi^s
\end{pmatrix},
\end{align}
where $p$ is the strength of the penalty factor which must be set large enough such that the constraints are satisfied under minimization. The objective function can be represented as a QUBO Hamiltonian
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
Q_{xx} & Q_{xs}\\
Q_{sx} & Q_{ss} 
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + pb^2,\\
\equiv & \Psi^T Q \Psi + C,
\label{eq:matrix_form}
\end{align}
where
\begin{align}
Q_{xx} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x + \mathrm{Diag}_{\psi^x}(c) T^x,\\
Q_{xs} = & Q_{sx}^T = p {T^{x}}^T A^T T^s,\\
Q_{ss} = & p\left[ {T^{s}}^T T^s + \mathrm{Diag}_{\psi^s}\left( {T^{s}}^T b + b^T T^s\right) \right].
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

@@

Given a graph $G(E,V)$, defined by the set of $V$ vertices and $E$ edges, and a subset of vertices $D \subseteq V$. If the set of all vertices not in $D$ is adjacent to at least one vertex in $D$, then $D$ is a dominating set. This is equivalent to requiring the set of nearest-neighbor vertices of $D$ (exclusive) and $D$ cover all vertices $N(D) \cup D = V$ (an example is given by Fig.~\ref{fig:dominating_sets}a). The set $D$ is a minimal dominating set if there is no proper subset of $D$ that is a dominating set, {\it{i.e.}}, the removal of any vertex in $D$ results in $N(D) \cup D  \neq V$. An example is given by Fig.~\ref{fig:dominating_sets}b. The domination number of $D$ is given by the cardinality of $|D| \equiv \overline{\overline{D}}$. The minimum dominating set (MDS) is defined by $D$ with the smallest domination number. Fig.~\ref{fig:dominating_sets}c shows an example of the minimum dominating set of $G(V, E)$ and is different from the minimal dominating set. We emphasize that while the maximum independent set is always a minimal dominating set as exemplified by Fig.~\ref{fig:dominating_sets}b, the minimum dominating set in general can have a smaller domination number. As a result, the solution to the dominating set problem can not be obtained by solving for the maximum independent set, a problem that is well studied for quantum annealers~\cite{}.

@@

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.
We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.
\begin{figure*}
	\centering
	\begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
	\end{tabular}
	\caption{
		Example of greedy MDS algorithm:
		The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
		Connected nodes are neighbors of the DS (turquoise) or already in the DS.
		In (a), one finds three unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
		In the first step $a\to b$, randomly the node (4) is picked.
		In (b), one has two unconnected choices [(1), (2)] with each adjacent to another unconnected node.
		After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
		The found solution is a (minimal) dominating set but not a MDS.
		To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
	}
	\label{fig:mds-greedy}
\end{figure*}

@@

The target Hamiltonian can be written in the matrix form
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
\mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] & - p(\mathbbm{1}+J^T)T^s\\
- p{T^s}^T(\mathbbm{1}+J)& p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + p \overline{\overline{V}}
\label{eq:matrix_form}
\end{align}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.
